Unsupervised Learning for Identifying Events in Active Target Experiments

### 2020-06-17 doc/Manuscript/article.tex
Updates from Overleaf

\title{Machine Learning Approaches  for Classification of Events in Active-Target Time Projection Chamber Experiments}

@@

The AT-TPC data we studied for this work was recorded as charge time-series for each of the  the $\sim10^4$ detector pads.
In this representation, an event is a record of 512 time-buckets for each of the detector pads. In our analysis, we represent each event as a downsampled 2D projection.
We chose to represent the data in 2D to facilitate the use of sophisticated image-recognition machine learning models.
The 2D representation of events was computed by viewing the event as a point cloud in the charge time-series and summing over the time-axis.
Additionally, we log-transform and normalize the data to the interval $[0, 1]$ for each point {\color{blue} are you sure we log transformed? were you given peaks data or the 2D projections? -MPK}. The final representation for each event is a $128 \times 128$ matrix with each element ranging from $0$ to $1$.

@@

\subsection{Why machine learning}

\begin{itemize}
    \item Traditional MC methods fall short in two principal ways:
    \begin{itemize}
        \item The computational cost per event is too large given the size of the data-sets
        \item The broken tracks, noisy environment creates bad fit statistics for otherwise useful events. End result is $f1 \sim 0.7 $
        \item define a ``bad" fit as a non-proton event
    \end{itemize}
\end{itemize}

@@

The MIXAE model comprises several autoencoders, each of which corresponds to a cluster. The number of autoencoders, and thus the number of clusters, has to be determined beforehand. Each autoencoder constructs a latent representation of a given example. Those representations are used as inputs to an auxiliary network which assigns probabilities to the clusters, indicating the likelihood that the given example belongs to each cluster. Examples are then assigned to the cluster with the highest probability. The models' objective can then be de-composed in three parts:
\begin{enumerate}
    \item Each batch should contain different classes
    \item Each predicted cluster assignment should be confident. 
    \item The reconstructions should be good. But as they are weighted by the cluster assignment, the importance of the reconstruction is proportional to the confidence in the cluster predictions.
\end{enumerate}

{\color{cyan} [Ryan: A description of the objective functions may or may not be worth including here, but I don't know what they are.]} The architecture is portrayed in Figure \ref{fig:mixae}, wherein tapered boxes denote a direction of compression in the network components.

@@

\noindent {\color{cyan} [Ryan: Still needs a concluding sentence or two here.]} For a more formal consideration of these heuristics, the model objective aand the assumptions made on the data by this model see \citet{Zhang}.

@@

\begin{figure}[H]
\centering
	\includegraphics[width=\textwidth]{custom_work/Filtered_mixae_conf_mat.pdf}
\caption{}\label{fig:mixae_confmat_filtered}
\end{figure}

\begin{figure}[H]
\centering
	\includegraphics[width=\textwidth]{custom_work/Raw_mixae_conf_mat.pdf}
\caption[MIXAE - confusion matrices]{Confusion matrices for the MIXAE clustering algorithm on raw AT-TPC events. The true labels indicate samples belonging to the p (proton), carbon (C), or other classes. We observe that in }\label{fig:mixae_confmat_raw}
\end{figure}

>>>>

\title{Unsupervised Learning  for Identifying  Events in Active Target  Experiments}

@@

The AT-TPC data we studied for this work was recorded as charge time-series for each of the  the $\sim10^4$ detector pads.
In this representation, an event is a record of $512$ time-buckets for each of the $10^4$ detector pads. In our analysis, we represent each event as a downsampled 2D projection.
We chose to represent the data in 2D to facilitate the use of sophisticated image-recognition machine learning models, and this data representation was shown to successfully classify the events of this experiment in a supervised setting \cite{Kuchera2019}.
First, the time-series data was represented as a 3D point cloud, where each point contains the maximum charge in the time-series trace. We log-transform to map the charge to the interval $[0, 1]$ for each point, then standardize the data. in the charge time-series and summing over the time-axis.
%
Additionally, we log-transform and normalize the data to the interval $[0, 1]$ for each point. The final representation for each event is a $128 \times 128$ matrix with each element ranging from $0$ to $1$.

@@

\subsection{Why machine learning}
{\bf we need to rewrite these bullet points, MHJ}
\begin{itemize}
    \item Traditional MC methods fall short in two principal ways:
    \begin{itemize}
        \item The computational cost per event is too large given the size of the data-sets
        \item The broken tracks with noisy environment create bad fit statistics for otherwise useful events. End result is f$_1 \sim 0.7 $
        \item define a ``bad" fit as a non-proton event
    \end{itemize}
\end{itemize}

@@

The MIXAE model comprises several autoencoders, each of which corresponds to a cluster. The number of autoencoders, and thus the number of clusters, has to be determined beforehand. Each autoencoder constructs a latent representation of a given example. Those representations are used as inputs to an auxiliary network which assigns scores to the clusters, indicating the likelihood that the given example belongs to each cluster. Examples are then assigned to the cluster with the highest score. The architecture is portrayed in Figure \ref{fig:mixae}, wherein tapered boxes denote a direction of compression in the network components.

@@

\noindent The objective during training is to enforce that these assumptions are met. For a more formal consideration of these heuristics, the model objective aand the assumptions made on the data by this model see \citet{Zhang}.

@@

\begin{figure}
\centering
	\includegraphics[width=\textwidth]{custom_work/Filtered_mixae_conf_mat.pdf}
\caption{}\label{fig:mixae_confmat_filtered}
\end{figure}

\begin{figure}
\centering
	\includegraphics[width=\textwidth]{custom_work/Raw_mixae_conf_mat.pdf}
\caption[MIXAE - confusion matrices]{Confusion matrices for the MIXAE clustering algorithm on raw AT-TPC events. The true labels indicate samples belonging to the p (proton), carbon (C), or other classes. We observe that in }\label{fig:mixae_confmat_raw}
\end{figure}
